---
title: "exercise_3"
author: "Cyril Geistlich"
date: "2023-05-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library("readr") 
library("sf") 
library("ggplot2") # to visualize data
library("gitcreds")
library("dplyr")
library("SimilarityMeasures")
```

```{r prep data}
boar<- read_delim("data/wildschwein_BE_2056.csv", ",")
boar$TierName |> unique()



sabi <- boar |> 
  filter(TierName == 'Sabi') |>
  filter(DatetimeUTC >= as.POSIXct("2015-07-01", tz = "UTC"), DatetimeUTC < as.POSIXct("2015-07-03", tz = "UTC"))
```

```{r visualize 1}
ggplot(sabi,aes(E,N, color = DatetimeUTC)) +
  geom_point() +
  geom_path() + 
  coord_equal()
```
```{r sampling interval}

sabi |>
  head(50) |>
  ggplot(aes(DatetimeUTC, 1)) +
  geom_point() +
  ggtitle("Sampling Interval Sabi")

```
```{r eucl distance}
sabi |> 
  mutate(
    stepLength = sqrt((lead(E)-E)**2 + (lead(N)-N)**2)
    )
```   
```{r}
sabi <- sabi |>
    mutate(
        nMinus2 = sqrt((lag(E, 2) - E)^2 + (lag(N, 2) - N)^2), 
        # distance to pos -30 minutes
        nMinus1 = sqrt((lag(E, 1) - E)^2 + (lag(N, 1) - N)^2), 
        # distance to pos -15 minutes
        nPlus1  = sqrt((E - lead(E, 1))^2 + (N - lead(N, 1))^2), 
        # distance to pos +15 mintues
        nPlus2  = sqrt((E - lead(E, 2))^2 + (N - lead(N, 2))^2) 
        # distance to pos +30 minutes
    )
```

```{r}
sabi <- sabi |>
    rowwise() |>
    mutate(
        stepMean = mean(c(nMinus2, nMinus1, nPlus1, nPlus2))
    ) |>
    ungroup()

sabi
```

```{r remove static points}
sabi <- sabi |>
    ungroup() |>
    mutate(static = stepMean < mean(stepMean, na.rm = TRUE))

sabi_filter <- sabi |>
    filter(!static)

sabi_filter |>
    ggplot(aes(E, N, color = DatetimeUTC)) +
    geom_path() +
    geom_point() +
    coord_fixed()
```
```{r}
ggplot(sabi, aes(stepMean)) +
  geom_histogram(binwidth = 10)
```
```{r vis static points}
sabi |>
  mutate(
    static = stepMean < mean(stepMean, na.rm = TRUE)
  ) |> 
  ggplot(aes(E,N)) +
  geom_point(aes(color = static)) +
  geom_path() +
  coord_fixed()
```

```{r read posmo}
posmo <- read_delim("data/posmo_2023-04-28T00_00_00+02_00-2023-04-30T23_59_59+02_00.csv", ",")
posmo <-  select(posmo, datetime, lon_x, lat_y)
posmo <- st_as_sf(posmo,coords = c("lon_x","lat_y"), crs = 4326) 
posmo <- st_transform(posmo, crs = 2056)
```

```{r prep data}
posmo_coordinates <- st_coordinates(posmo)
posmo <- cbind(posmo, posmo_coordinates)
posmo <- posmo |>
    filter(as.Date(datetime) == "2023-04-29")
```

# Task 1: Segmentation

```{r sampling interval Posmo}
posmo |>
  ggplot(aes(datetime, 1)) +
  geom_point() +
  ggtitle("Sampling Interval Posmo")

```

# Task 2: Specify and apply threshold distance

```{r steplenght Posmo}
posmo |> 
  mutate(
    stepLength = sqrt((lead(X)-X)**2 + (lead(Y)-Y)**2)
    )
```
```{r prep stepmean}
posmo <- posmo |>
    mutate(
        nMinus2 = sqrt((lag(X, 2) - X)^2 + (lag(Y, 2) - Y)^2), 
        nMinus1 = sqrt((lag(X, 1) - X)^2 + (lag(Y, 1) - Y)^2), 
        nPlus1  = sqrt((X - lead(X, 1))^2 + (Y - lead(Y, 1))^2), 
        nPlus2  = sqrt((X - lead(X, 2))^2 + (Y - lead(Y, 2))^2) 
    )
```

```{r calc stepmean}
posmo <- posmo |>
    rowwise() |>
    mutate(
        stepMean = mean(c(nMinus2, nMinus1, nPlus1, nPlus2))
    ) |>
    ungroup()


posmo <- tidyr::drop_na(posmo)

posmo <- posmo |>
  ungroup() |>
  mutate(static = stepMean < mean(stepMean, na.rm = TRUE))

posmo_mean <- posmo |>
  filter(!static) 

```

```{r visualise stepmean}
ggplot(posmo, aes(stepMean)) +
  geom_histogram(binwidth = 10)
  
```

# Task 3: Visualise segmented trajectories

```{r visualise static}
posmo |>
    ggplot(aes(X, Y, color = static)) +
    geom_path() +
    geom_point() +
    coord_fixed() +
    ggtitle("False Flag stepMean < mean(stepmean)")

posmo_mean|>
    ggplot(aes(X, Y, color = datetime)) +
    geom_path() +
    geom_point() +
    coord_fixed()

  table(posmo$static)
```
# Task 4: Segment-based analysis

```{r assign id}


rle_id <- function(vec) {
    x <- rle(vec)$lengths
    as.factor(rep(seq_along(x), times = x))
}

rle_id(posmo$static)
```

```{r rename and filter static}

posmo <- posmo |>
    mutate(segment_id = as.numeric(rle_id(static))) |>
    filter(!static) |>
    mutate(diff_s = as.numeric(difftime(lead(datetime),datetime)))

#Remove last time from each segment
posmo$diff_s_rm_last <- ave(posmo$diff_s, posmo$segment_id, FUN = function(x) c(x[-length(x)], NA))

# sum the 'diff_s_without_last' column for each segment_id
tot_diff_s <- aggregate(posmo$diff_s_rm_last, by=list(segment=posmo$segment_id), FUN=sum, na.rm=TRUE)

# rename the columns to something more descriptive
colnames(tot_diff_s) <- c("segment_id", "tot_diff_s")


posmo <- left_join(posmo, tot_diff_s, by = "segment_id")

```

```{r visualise}
posmo |>
  mutate(segment_id = as.factor(segment_id))|>
    ggplot(aes(X, Y, color = segment_id)) +
    geom_path()+
    geom_point() +
    coord_fixed() + 
    ggtitle("Segmentation (All)") +
    scale_color_discrete(name = "Segment ID")


posmo |>
  mutate(segment_id = as.factor(segment_id))|>
  filter(tot_diff_s > 300) |>
  ggplot(aes(X, Y, color = segment_id)) +
    geom_path() +
    geom_point() +
    coord_fixed() + 
    ggtitle("Segmentation (> 5 minutes)") +
    scale_color_discrete(name = "Segment ID")
```
```{r clear env}
rm(list = ls())
```

# Task 5: Similarity Measures

```{r}
pedestrian <- read_delim("data/pedestrian.csv", ",")
```

```{r visualise pedestrians}
pedestrian |> 
  ggplot(aes(E,N,colour = as.factor(TrajID))) +
  geom_point() +
  geom_path() +
  facet_wrap(~ TrajID, ncol = 3) +
  coord_fixed() + 
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
  

```
Before visualizing your results think about the following: Which two trajectories to you percieve to be most similar, which are most dissimilar? Now visualize the results from the computed similarity measures. Which measure reflects your own intuition the closest?

The paths are very similar overall, especially 1,2,3 & 6. Path 4 is slightly different and path 5 might have an outlier. I expect the similarity measures to reflect these results.  

# Task 6: Calculate Similarity

```{r Compute similarity}
# Select the first trajectory
traj1 <- filter(pedestrian, TrajID == 1) |> select(E, N) |> as.matrix()

# Create storage for similarity measures
similarity_measures <- c()

# Compute similarity measures
for (trajID in unique(pedestrian$TrajID)[-1]) {
  traj2 <- filter(pedestrian, TrajID == trajID) |> select(E, N) |> as.matrix()
  sim <- c(DTW(traj1, traj2),
           EditDist(traj1, traj2, pointDistance = 20),
           Frechet(traj1, traj2, testLeash = -1),
           LCSS(traj1, traj2, pointSpacing = 5, pointDistance = 20, errorMarg = 5, returnTrans = FALSE))
  similarity_measures <- rbind(similarity_measures, sim)
}
```

```{r}
(similarity_measures)
```

```{r}
# Add similarity measures to the pedestrian data frame
pedestrian_similarity <- data.frame(TrajID = unique(pedestrian$TrajID)[-1],
                                     DTW = similarity_measures[, 1],
                                     EditDist = similarity_measures[, 2],
                                     Frechet = similarity_measures[, 3],
                                     LCSS = similarity_measures[, 4])
write.csv(pedestrian_similarity, "data/similarities.csv", row.names = FALSE)
```

```{r results similarity}
# Results
pedestrian_similarity
```

```{r vis results}
pedestrian_similarity_long <- pedestrian_similarity |> 
  tidyr::pivot_longer(cols = c(DTW, EditDist, Frechet, LCSS), names_to = "method")

ggplot(pedestrian_similarity_long, aes(x = TrajID, y = value, fill = as.factor(TrajID))) +
  geom_col() +
  facet_wrap(~method, nrow = 2, scales = "free_y") +
  labs(x = "TrajID", y = "Similarity Value", fill = "TrajID") +
  theme(legend.position = "right")
  
```

The calculated results differ from the expectations. I expected path 4 & 5 to be very different. In fact they fall in between in the similarity measures. This is unexpected and an adjustment of input parameters could yield better results. 
